Advancements in Moltres for Time-Dependent Multiphysics Molten Salt Reactor Modeling

### 2025-01-03 abstract.tex
Edits before format review

\glspl{MSR} are advanced reactors noted for strong passive safety features.
They present unique challenges in multiphysics reactor modeling \& simulation arising from
strong temperature reactivity feedback, delayed neutron precursor flow, and turbulent heat transport
in fuel salt regions. Simulating the complex multiphysics interactions in
\glspl{MSR} requires robust, flexible, and highly scalable multiphysics software. Many \gls{MSR}
designs also still retain control rods which can complicate time-dependent reactivity-initiated
transient simulations on reactor software relying on neutron diffusion theory.
This work builds on existing capabilities in Moltres, a MOOSE-based \gls{MSR} simulation software,
to tackle these challenges and support efforts towards \gls{MSR} deployment.

This work verified and validated existing multiphysics coupling capabilities in Moltres in two
comparative studies. In both studies, Moltres showed good agreement with other \gls{MSR} simulation tools
involving coupled neutronics and thermal-hydraulics problems. This work also introduces a
turbulence model in Moltres to support future \gls{MSR} analyses involving turbulent delayed
neutron precursor and temperature transport. Lastly, this work introduces a novel hybrid
$S_N$-diffusion method for accurate control rod modeling in time-dependent \gls{MSR} simulations.
The hybrid method combines the strengths of both approaches by generating transport corrections
using the $S_N$ method near control rods. The $S_N$ and neutron diffusion solvers are coupled
through an adaptive boundary coupling algorithm. This algorithm allows the solver to adapt to the
transport correction parameters and preserve smooth neutron flux gradients across the interface.
In 1-D, 2-D, and 3-D $k$-eigenvalue simulations, the hybrid method produced accurate control
rod worth estimates, relative to reference neutron transport solutions and experimental data,
at approximately four times the computational cost of the neutron diffusion method.
%Demonstrations of the hybrid method for time-dependent rod drop and reactivity insertion
%simulations also performed well in reproducing expected trends observed in experimental data.
A demonstration of the hybrid method for a time-dependent rod drop
simulation also performed well in reproducing expected trends observed in experimental data.
Analysis of its computational performance indicated the possibility of further optimizations
beyond its current implementation. With the hybrid method's spatial resolution and
efficient computational performance, Moltres could enable accurate and cost-effective simulations
of asymmetric transients in \glspl{MSR}.

>>>>

\glspl{MSR} are advanced reactors noted for strong passive safety features.
They present unique challenges in multiphysics reactor modeling \& simulation arising from
strong temperature reactivity feedback, delayed neutron precursor flow, and turbulent heat transport
in fuel salt regions. Simulating the complex multiphysics interactions in
\glspl{MSR} requires robust, flexible, and highly scalable multiphysics software. Many \gls{MSR}
designs also still retain control rods which can complicate time-dependent reactivity-initiated
transient simulations on reactor software relying on neutron diffusion theory.
This work builds on existing capabilities in Moltres, a MOOSE-based \gls{MSR} simulation software,
to address these challenges and support efforts towards \gls{MSR} deployment.

This work verified and validated existing multiphysics coupling capabilities in Moltres in two
comparative studies. In both studies, Moltres showed good agreement with other \gls{MSR} simulation tools
involving coupled neutronics and thermal-hydraulics problems. This work also introduces a
turbulence model in Moltres to support future \gls{MSR} analyses involving turbulent delayed
neutron precursor and temperature transport. Lastly, this work introduces a novel hybrid
$S_N$-diffusion method for accurate control rod modeling in time-dependent \gls{MSR} simulations.
The hybrid method combines the strengths of both approaches by generating transport corrections
using the $S_N$ method near control rods. The $S_N$ and neutron diffusion solvers are coupled
through an adaptive boundary coupling algorithm. This algorithm allows the solver to adapt to the
transport correction parameters and preserve smooth neutron flux gradients across the interface.
In 1-D, 2-D, and 3-D $k$-eigenvalue simulations, the hybrid method produced accurate control
rod worth estimates, relative to reference neutron transport solutions and experimental data,
at approximately four times the computational cost of the neutron diffusion method.
%Demonstrations of the hybrid method for time-dependent rod drop and reactivity insertion
%simulations also performed well in reproducing expected trends observed in experimental data.
The hybrid method also performed well in a time-dependent rod drop simulation
by reproducing expected trends observed in experimental data.
Analysis of its computational performance indicated the possibility of further optimizations
beyond its current implementation. With the hybrid method's spatial resolution and
efficient computational performance, Moltres could enable accurate and cost-effective simulations
of asymmetric transients in \glspl{MSR}.

### 2025-01-03 appendix.tex
Full draft + edits
Edits before format review

\chapter{Numerical Implementation of Neutronics Solvers on Python} \label{chap:implementation}

@@

>>>>

\chapter{Numerical Implementation of Neutronics Solvers in Python} \label{chap:implementation}

@@

\chapter{3-D \gls{MSRE} Neutron Flux Distributions}

This appendix contains figures showing the group 1 and 8 neutron flux distributions in the 3-D
\gls{MSRE} full-core model with Rod 1 inserted by 27.5 and 51 inches.

\begin{figure}[p]
  \centering
  \includegraphics[width=\columnwidth]{halfway-g1}
  \caption{Group 1 neutron flux distribution with Rod 1 inserted by 27.5 inches.}
  \label{fig:halfway-g1}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=\columnwidth]{halfway-g8}
  \caption{Group 8 neutron flux distribution with Rod 1 inserted by 27.5 inches.}
  \label{fig:halfway-g8}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=\columnwidth]{inserted-g1}
  \caption{Group 1 neutron flux distribution with Rod 1 inserted by 51 inches.}
  \label{fig:inserted-g1}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=\columnwidth]{inserted-g8}
  \caption{Group 8 neutron flux distribution with Rod 1 inserted by 51 inches.}
  \label{fig:inserted-g8}
\end{figure}

### 2025-01-03 conclusion.tex
Edits before format review

Chapters \ref{chap:msre} and \ref{chap:transient} demonstrated the hybrid $S_N$-diffusion method's
accuracy in control rod worth calculations. However, some discrepancies still persist in the
$k_\text{eff}$ estimates of individual reactor states. These discrepancies affect the average
temperature of the reactor in multiphysics simulations to compensate for the mismatch in
reactivity. These discrepancies could be reduced through
neutron leakage correction at the vacuum boundaries. Remedies for this issue exist in a number of
forms such as group-wise or matrix albedo boundary conditions.

The 3-D simulations in this work uncovered significant memory usage by the $S_N$ solver, even
with the distributed mesh feature to spread mesh and variable storage across the compute nodes.
Each full-core simulation required at least 40 nodes with 512 GB memory each to run.
This issue may be a stumbling block for using the hybrid method on smaller computing clusters that
have less memory per processor. A custom preconditioner and solver routine in Moltres could help to
reduce memory usage by referencing the same stored Jacobian C++ variable for neutron angular flux
variables in the same neutron energy group and on the same \gls{FEM} quadrature point. Their
Jacobian formulations are identical aside from the level-symmetric ordinate and weight. Another
area for optimization in the hybrid method is reducing data transfer times between the $S_N$ and
diffusion solvers. Due to mesh distribution, each 3-D simulation spent 18 \% of its solution time
on data transfers between processors. This could be minimized through optimizations in the mesh
distribution and the data transfer caching systems.

>>>>

Chapters \ref{chap:msre} and \ref{chap:transient} demonstrated the hybrid $S_N$-diffusion method's
accuracy in control rod worth calculations. However, some discrepancies still persist in the
$k_\text{eff}$ estimates of individual reactor states. These discrepancies affect the average
temperature of the reactor in multiphysics simulations to compensate for the mismatch in
reactivity. These discrepancies could be reduced through
neutron leakage correction at the vacuum boundaries. Remedies for the $k_\text{eff}$ mismatch
exist in a number of forms such as group-wise or matrix albedo boundary conditions.

The 3-D simulations in this work uncovered significant memory usage by the $S_N$ solver, even
with the distributed mesh feature to spread mesh and variable data storage across the compute nodes.
Each full-core simulation required at least 40 nodes with 512 GB memory each to run.
This issue may be a stumbling block for using the hybrid method on smaller computing clusters that
have less memory per processor. A custom preconditioner and solver routine in Moltres could help to
reduce memory usage by referencing the same stored Jacobian C++ variable for neutron angular flux
variables in the same neutron energy group and on the same \gls{FEM} quadrature point. Their
Jacobian formulations are identical aside from the level-symmetric ordinate and weight. Another
area for optimization in the hybrid method is reducing data transfer times between the $S_N$ and
diffusion solvers. Due to mesh distribution, each 3-D simulation spent 18 \% of its solution time
on data transfers between processors. This could be minimized through optimizations in the mesh
distribution and the data transfer caching systems.

Time-dependent simulations in this work uncovered two issues: rod cusping effects when a moving
control rod does not align with the mesh interfaces, and slow solution convergence rates during
control rod motion. While this work applied an empirical technique to correct for rod cusping
effects, Moltres would benefit from a more robust technique for the long term. Slow solution
convergence rates could be resolved by investigating better nested solver coupling structures to
mitigate the effects of lagged solutions and applying solution relaxation schemes.
